---
title: "index"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
THe Spotify dataset by year contains data from 1922 to 2021, collected by Yamac Eren using Spotify's web API. The dataset contains information on acousticness, danceability, duration, etc.

```{r include=FALSE}
# Source in the other scripts
source("scripts/summary_table.R")
source("scripts/popularity_by_year.R")
source("scripts/dance_instrumental.R")

# Global Data
df = read.csv("data/Spotify/data_by_year_o.csv")

# Loading `knitr` library
library(knitr)

```
## Summary Table
Below shows a table on the progression of music throughout the decades.
It displays the averages for various categories of all music for each respective
decade. This table shows us that over the decades, music has gotten less 
acoustic, instrumental, and live. But, it has increased in danceability, energy,
loudness, and tempo.
```{r echo=FALSE}
kable(table_summary(df))
```

Below you will find different visualizations that show how music has played a role in U.S. history.

### Popularity by Year

Below you will find a chart that shows a scatter plot visualization of how 
music popularity has changed throughout the years. 

```{r, echo=FALSE}
popularity_by_year(df)
```

We choose to use a scatter plot to show data because it best demonstrated the 
trend in music popularity. We found that the overall music popularity has seen 
a continuous increase over the years and has continued to become a more 
prominent part of our culture. The scatter plot shows a strong positive 
correlation between the popularity of music (y-axis) over the years (x-axis) 
and we included a line of best fit to show the general trend. 

### Danceability and Instrumentalness by Year

```{r, echo=FALSE}
inst_dance_chart(df)
```

We chose to use a stacked bar chart as the intent was to compare the overall change in "danceability" and "instrumentalness" of music over time, to see if there exist a relationship between "danceability" and "instrumentalness". Based on the chart, we found that from 1921 to 2020, there seems to be no significant change in "danceability". However, there is clearly a downward trend in "instrumentalness".
