---
title: "index"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
THe Spotify dataset by year contains data from 1922 to 2021, collected by Yamac Eren using Spotify's web API. The dataset contains information on acousticness, danceability, duration, etc.





```{r include=FALSE}
# Source in the other script
source("scripts/popularity_by_year.R")

# Global Data
df = read.csv("data/Spotify/data_by_year_o.csv")

# Loading `knitr` library
library(knitr)

```
Below you will find different visualizations that show how music has played a
role in U.S. history.

### Popularity by Year

Below you will find a chart that shows a scatter plot visualization of how 
music popularity has changed throughout the years. 

```{r, echo=FALSE}
popularity_by_year(df)
```

We choose to use a scatter plot to show data because it best demonstrated the 
trend in music popularity. We found that the overall music popularity has seen 
a continuous increase over the years and has continued to become a more 
prominent part of our culture. The scatter plot shows a strong positive 
correlation between the popularity of music (y-axis) over the years (x-axis) 
and we included a line of best fit to show the general trend. 

