---
title: "Analysis on Audio Features of Songs From 1921 to 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
The Spotify dataset by year contains data from 1921 to 2020, collected by Yamac Eren using Spotify's web API. The dataset consists of six different csv files and a json file. 

```{r include=FALSE}
# Source in the other scripts
source("scripts/summary_table.R")
source("scripts/popularity_by_year.R")
source("scripts/dance_instrumental.R")
source("scripts/loudness_by_year.R")
source("scripts/summary_info.R")

# Global Data
df = read.csv("data/Spotify/data_by_year_o.csv")

# Loading `knitr` library
library(knitr)
library("tidyverse")
library("ggplot2")
library("dplyr")
library("forestmangr")

```

### Summary Information

```{r, include=FALSE}
summary <- get_summary_info(df)
```

>When using this data set there are a few things which are both interesting 
and beneficial to note. First, the number of rows in the data frame is 
`r summary[3]` which each correlate to the years from 1921-2020. Throughout the 
data frame, the average length of songs this century is `r summary[4]` minutes. 
The most common key used throughout the century was the key of `r summary[2]`. 
Interestingly, the year in which music was the least loud was `r summary[5]`. 
Finally, the highest amount of danceability recorded in the data set was 
`r summary[1]`.

### Summary Table
>Below shows a table on the progression of music throughout the decades.
It displays the averages of various categories for music in its respective 
decade. For example, in the 1970s, 40% of all music was acoustic. This table shows us 
that over the decades, music has gotten less acoustic, instrumental, and live. 
But, it has increased in danceability, energy, loudness, and tempo.

```{r echo=FALSE}
kable(table_summary(df))
```


## Data Visualization of Audio Features and Popularity of Music

### Popularity by Year

Below you will find a chart that shows a scatter plot visualization of how 
music popularity has changed throughout the years. 

```{r, echo=FALSE, message = FALSE}
popularity_by_year(df)
```

>We chose to use a scatter plot to show data because it best demonstrated the 
trend in music popularity. We found that the overall music popularity has seen 
a continuous increase over the years and has continued to become a more 
prominent part of our culture. The scatter plot shows a strong positive 
correlation between the popularity of music (y-axis) over the years (x-axis) 
and we included a line of best fit to show the general trend. 

### Danceability and Instrumentalness by Year

Below you will find a chart that shows a bar representation of how 
the danceability and instrumentalness features have changed over the years. 

```{r, echo=FALSE}
inst_dance_chart(df)
```

>We chose to use a stacked bar chart as the intent was to compare the overall change in "danceability" and "instrumentalness" of music over time, to see if there exist a relationship between "danceability" and "instrumentalness". Based on the chart, we found that from 1921 to 2020, there seems to be no significant change in "danceability". However, there is clearly a downward trend in "instrumentalness".

### Loudness by Year

Below you will find a chart that shows a linear representation of how 
the loudness features (dB) has changed over the years.

```{r, echo=FALSE, message = FALSE}
plot_loudness_year(df)
```

>We chose to use a line graph with a smooth line to clearly visualize the change
in the loudness of music (dB) over time. The smoothline further aids the audience
to capture the overall trend of the loudness level of music from 1921 to 2020.
From our observation, we discovered that the change in loudness level from the early
1920s to 1950s was very dynamic, ranging from -19.275dB to -13.864. Also, the smooth
line conveys that the overall loudness level of music has been gradually increasing,
reaching its maximum level of -6.595067dB  in year 2020.